Technical Report

### 2017-11-07 fully-abstract-comp/technical-report.tex
corrected where the values from the stack pointer comes from.
Typos

\dominique{27-09-2017: I think we need to do something about tail calls in the source language?}
\lau{27-09-2017: In what sense? Do you want to allow them (I do not thinkn they are allowed at the moment?)}

@@

The $\decInstr{}$ function must be surjective and injective for all non-$\tfail$ instructions. The $\encInstr{}$ function must be injective. Futher $\decInstr{}$ must be the left inverse of $\encInstr{}$ that is for all $i \in \Instr$
\[
  \decInstr{}(\encInstr{i}) = i
\]

@@

\paragraph{Call implementation}
\label{par:call-impl}
\lau{16-10-2017: Fill this paragraph with $\scall{}{}$ implementation, i.e., the instructions $\scall[1]{r_1}{r_2} \dots \scall[\calllen]{r_1}{r_2}$ is equal to. Add convenience notation like $\update{\mem.a}{\scall{r_1}{r_2}} = \dots$}
\lau{16-10-2017: Implementation of $\scall{}{}$ may need registers for temporary registers - we should update the semantics accordingly.}
\[
  \begin{array}{l}
    \tgete{r_{1t}}{\rstk}\\
    \tstore{\rstk}{r_{t1}}\\
    \tcca{\rstk}{-1}\\
    \tgeta{r_{t1}}{\rstk}\\
    \tsplit{\rstk}{\rretd}{\rstk}{r_{t1}}\\
    \tmove{r_{t1}}{\pcreg}\\
    \tsetatob{r_{t1}}{r_{t1}}\\
    \tload{r_{t1}}{r_{t1}}\\
    \tcca{r_{t1}}{{\color{red} \dots}}\\
    \tcseal{\rretd}{\rretd}{r_{t1}}\\
    \tmove{\rretc}{\pcreg}\\
    \tcca{\rretc}{5} \text{ //magic number is offset to return code}\\
    \tcseal{\rretc}{\rretc}{r_{t1}}\\
    \tmove{r_{t1}}{0}\\
    \txjmp{r_1}{r_2}\\
    \text{// the following is the return code}\\
    \tgetb{r_{t1}}{\rstk}\\
    \tminus{r_{t1}}{t_{t1}}{\stkb} \text{ //$\stkb$ is the stack base constant} \\
    \tmove{r_{t2}}{\pcreg}\\
    \tcca{r_{t2}}{5} \text{ //magic number is the offset to fail} \\
    \tjnz{r_{t2}}{r_{t1}} \\
    \tcca{r_{t2}}{1} \text{ //magic number is the offset to after fail} \\
    \tjmp{r_{t2}} \\
    \tfail \\
    \tsplice{\rstk}{\rstk}{\rdata} \\
    \tcca{\rstk}{1}\\
    \tmove{r_{t1}}{0}\\
    \tmove{r_{t2}}{0}
  \end{array}
\]
{\color{red} \dots} needs to be the offset to the seal we want to use.
The call code does the following:
\begin{itemize}
\item Store the end address of the stack to the stack (this ensures the stack is non-empty), and decrement the pointer according to convention.
\item Get the current address of the stack pointer and split according to it.
\item Retrieve the seal of the program.
\item cca the seal, so the seal to be used is active 
\item seal our private part of the stack capability.
\item Move the pc out of the pc register, adjust it to point to the first address of the return code, and seal it.
\item clear the temporaryu register.
\item cross jump to the two specified registers.
\item Upon return:
  \begin{itemize}
  \item get the base of the stack and check that it matches up with the global base of the stack.If not, fail.
  \item splice the returned stack pointer and the stack pointer for our private stack.
  \item adjust stack pointer to first empty address (the address with the end address is considered free)  
  \item clear the temporary registers.
  \end{itemize}
\end{itemize}

@@

\begin{definition}
  \label{def:wxorx}
  ~\\
  A word $w$ is said to respect write-xor-execute iff
  \begin{itemize}
  \item $w \in \ints$
  \item $w = \seal{\_,\_,\_,\_}$
  \item $w = ((\perm,\_),\_,\_,\_) \tand \perm \neq \rwx$
  \item $\seal{\_,\vsc}$ and $\vsc$ respects write-xor-execute
  \item $\src{\stkptr{\perm} \tand \perm \neq \rwx}$
  \item $\src{w = \retptrc(\_,\_)}$
  \item $\src{w = \retptrd(\_,\_)}$
  \end{itemize}
~\\
  A register file $\reg$ is said to respect write-xor-execute with slack $[\bar{r}]$ iff \\
  For all $r \in RegName$ $\reg(r) \setminus [\bar{r}]$ respects write-xor-execute.
~\\
~\\
  A memory segment $\ms$ is said to respect write-xor-execute iff\\
  For all $\aaddr \in \dom(\ms)$ $\ms(\aaddr)$ respects write-xor-execute
\end{definition}

@@

\section{Logical Relation}
\subsection{Worlds}
\lau{30-10-2017: TODO: solve recursiveness in domain equation in the following. }
\lau{02-11-2017: We may need revoked regions again to handle the stack.}
\lau{06-11-2017: If we need to introduce revoked regions like we hade for the local capabilities, we may consider the following: For the future world relation require the existance of a bijection from region names to region names. For all regions in the present world the future region is given by the bijection. This would mean that reinstantiation would happen simply by giving a bijection that maps to the new region. In other words, the weird shadowing would not be necessary.
  \begin{mathpar}
    \inferrule{ \dom(W') \supseteq \dom(W) \\ \exists m : \RegionName \fun \RegionName, \text{ $m$ injective} \ldotp \forall r \in \dom(W) \ldotp W'(m(r)) \pubft W(r) }
    { W' \pubft W }
  \end{mathpar}
}
\[
  \World = \RegionName \parfun \Region
\]
where $\RegionName = \nats$.
\[
  \Region = \left\{
  \begin{array}{l}
    \{\spatial \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})\\
    \{\spatialo \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})\\
    \{\pure \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})
  \end{array} \right.
\]
where $\spatial$ and $\spatialo$ are regions governing segments of memory addressed by linear capabilities. $\spatialo$ signifies that this region is addressable. $\spatial$ signifies that the region is not owned and can thus not be addressed. Finally, $\pure$ signifies that the region is only addressed by non-linear capabilities. Notice that no region allows for both linear and non-linear capabilities to address it.
\lau{26-20-2017: We may also want to build the infrastructure for the stack more explicitely into the worlds (but maybe we can handle it in the memory satisfaction).}
\lau{26-20-2017: It is not clear to me what the monotonicity requirements on the memory interpretation functions should be.}
  \[
   \Rels = \{\phi_\pub \times \phi \in \powerset{\State^2}\times \powerset{\State^2} \mid \phi_\pub, \phi \text{ is reflexive and transitive and } \phi_\pub \subseteq \phi \}
  \]

Disjoint union of worlds. Joins together two alike worlds with strictly disjoint ownership over $\spatialo$-regions.
\[
  W_1 \oplus W_1 = W
  \begin{array}[t]{l}
    \text{if $\dom(W) = \dom(W_1) = \dom(W_2) \tand$} \\
    \forall r \in \dom(W) \ldotp W(r) = W_1(r) \oplus W_2(r)
  \end{array}
\]
$\oplus$ is defined as follows
\begin{align*}
  (\pure,\var{sm}) \oplus (\pure,\var{sm}) =  & \; (\pure,\var{sm}) \\
  (\spatialo,\var{sm}) \oplus (\spatial,\var{sm}) = & \; (\spatial,\var{sm}) \oplus (\spatialo,\var{sm})\\
                                           =  & \; (\spatialo,\var{sm}) \\
  (\spatial,\var{sm}) \oplus (\spatial,\var{sm}) =  & \; (\spatial,\var{sm}) \\
\end{align*}
and for all other cases $\oplus$ is undefined. Specifically, $\oplus$ is not defined when both sides are a $\spatialo$-region. It is further not defined if the two sides do not agree on region type and state machine. Here $\var{sm}$ is the state machine of the region.

@@

\subsection{Memory satisfaction}
\[
  \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W} \text{ iff } 
  \left\{
    \begin{array}{l}
      \exists R_\ms : \dom(W) \fun \MemSeg^2 \ldotp \\
      \quad \ms_T = \biguplus_{r \in \dom(W)} \pi_2(R_\ms(r)) \wedge \\
      \quad \stk = (\_,\ms_{\stk,0}) :: \dots :: (\_,\ms_{\stk,m}) \wedge \\
      \quad \exists r_0 , \dots, r_m, r_{m+1} \ldotp \pi_1(R_\ms(r_0,\dots,r_m)) = \ms_{\stk,0},\dots , \ms_{\stk,m} \wedge \\
      \qquad \pi_1(R_\ms(r_{m+1})) = \ms_\stk \wedge \\
      \qquad \ms_S \uplus \ms_\stk \uplus \ms_{\stk,0} \uplus \dots \ms_{\stk,m} = \biguplus_{r \in \dom(W)} \pi_1(R_\ms(r)) \wedge \\
      \qquad \exists R_w : \dom(W) \fun \World \ldotp\\
      \quad \qquad W = \bigoplus_{r \in \dom(W)} R_w(r) \wedge \\
      \quad \qquad \forall r \in \dom(W) , n' < n \ldotp \\
      \qquad \qquad \npair[n']{R_\ms(r)} \in W(r).H \; (W(r).s) \; (\xi^{-1} (R_w (r))) 
    \end{array}
  \right.
\]
In the definition of memory satisfaction, $R_\ms$ acts as a partitioning in the two related memories. As the source language is a special interpretation of the target language, part of the target memory needs to relate to the stack in the source language. We also need to split up the world which is done by $R_W$.We split up the world in order to make sure that the linearity constraints impossed by the world are satisfied.
\lau{30-10-2017: It is not clear to me whether this definition should also talk about $\opc$.}

\subsection{Relation}
\lau{19-10-2017: Rest of source configuration needs to come from somewhere in $\lre$.}
\lau{19-10-2017: Maybe linearity should be taken care of in just one place (instead of in $\lrr$ relation and mem staisfaction). That is, add a predicate that checks whether the configuration satisfies linearity (such a condition may be to syntactic to work?).}
In the following definitions, \src{blue} is used to indicate values related to the source machine. This is unlike previous definitions, where \src{blue} is used  indicate source language specific parts of definitions.
\[
  \lre(W) = \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
    \begin{array}{l}
      \forall n' \leq n, \src{k_{d,S}}, \src{k_{c,S}}, k_{d,T}, k_{c,T}, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
      \quad \forall W_\lrk, W_\lrr, W_\lrm \ldotp \\
      \qquad\npair[n']{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \in \lrp(W) \wedge \\
      \qquad\npair[n']{\stpair[.]{k_{d,S},k_{c,S}}{k_{d,T},k_{c,T}}} \in \lrk(W_\lrk) \wedge \\
      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
      \qquad\Rightarrow\\
      \qquad\npair[n']{\left(\array{l}\src{(\ms_S,\reg_S\updReg{\pcreg,\rdata,\rretc,\rretd}{v_{c,S},v_{d,S},k_{c,S},k_{d,S}},\stk, \ms_\stk)},\\
                                          (\ms_T,\reg_T\updReg{\pcreg,\rdata,\rretc,\rretd}{v_{c,T},v_{d
,T},k_{c,T},k_{d,T}})\endarray\right)}
      \in \lro(W\oplus W_\lrk \oplus W_\lrr \oplus W_\lrm)
    \end{array}
    \right\}
\]
\lau{24-10-2017: Dominique, I do not see how the splitting of worlds is going to work out. I can see how it will enforce linearity, but is you give a "spatial" region (the kind of region that is only allowed to go to one part of a split) to the $\lrr$-relation, then the memory satisfation does not have it, and it cannot find a memory segment for this part. In other words, the memory satisfaction will not be able to find a concrete memory segment for segments governed by a linear capability in the register file. Maybe the memory satisfaction should have all of the parts, so it can represent them (but also limit what linear capabilities can be used).}

@@

\lau{03-11-2017: TODO take a look at linearity of this definition - in paticular the recovered pc and stack pointer.}
\[
  \lrk(W) = \left\{ \npair{\left(\array{l}\src{\retptrd(\sigma,(\aaddr_d,\eaddr_d))},\\\src{\retptrc(\sigma,(\perm,\lin,\baddr_c,\eaddr_c,\aaddr_c))},\\k_{d,T},k_{c,T}\endarray\right)} \middle|
    \begin{array}{l}
      k_{d,T} = \sealed{\sigma,c_d} \wedge\\
      k_{c,T} = \sealed{\sigma,c_c} \wedge\\
      c_c = ((\perm,\lin),\baddr_c,\eaddr_c,\aaddr_c) \wedge\\
      c_d = ((\rw,\linear),\aaddr_d,\eaddr_d,\aaddr_d-1) \wedge\\
      \forall n' \leq n, W' \pubft W \ldotp \\
      \quad \forall, W_\lrr, W_\lrm \ldotp  \\
      \qquad\opc = c_c \wedge\\
      \qquad\dom(\ms_{\stk,f}) = [\aaddr_d,\eaddr_d] \wedge\\
      \qquad\npair[n']{\src{\retptrd(\sigma,\_)},\src{\retptrc(\sigma,\_)},k_{d,T},k_{c,T}} \in \lrp(W') \wedge \\
      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk::(\opc,\ms_{\stk,f}),\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
      \qquad\dom(\ms_\stk) = [\baddr_\stk,\eaddr_\stk] \wedge \\
      \qquad\eaddr_\stk = \aaddr_d-1\\
      \qquad\Rightarrow\\
      \qquad\npair[n']{\left(\array{l}\src{\left(\arraycolsep=0pt\array{l}\ms_S,\reg_S\updReg{\pcreg,\rstk}{\opc,\stkptr{\rw,\baddr_\stk,\eaddr_d,\_}},\\\stk, \ms_\stk \uplus \ms_{\stk,f}\endarray\right)},\\
                                          (\ms_T,\reg_T\updReg{\pcreg,\rstk}{\opc,((\rw,\lin),\baddr_\stk,\eaddr_d,\_)})\endarray\right)}
      \in \lro(W' \oplus W_\lrr \oplus W_\lrm)
    \end{array}
  \right\}
\]

@@

\[
  \lrr(W) = \left\{ \npair{\stpair{\reg}{\reg}} \middle|
    \begin{array}{l}
      \exists S : \RegName \fun \World \ldotp \\
      \quad S \text{ is a separation of $W$} \tand\\
      \quad \forall r \in \RegName \setminus \{\pcreg, \rretd,\rretc \}\ldotp\\
      \qquad\npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrv(S(r))
    \end{array}
            \right\}
\]
Where $S\text{ is a separation of $W$}$ means that no two partion share the same ``spatial'' region (whatever that is).

@@

One way to solve this problem is to make the "top addess" of the stack known. There are many ways to do this, but we have chosen to do the following:
\begin{itemize}
\item The stack grows downwards, so the ``last address'' of the stack is the base address of the initial stack capability.
\item The base address of the stack capability is a fixed address, so in the semantics, it will be expressed as a constant that is publicly known.
\item At some point before a call, it must be checked whether the stack we are using actually has the globally known base address (if not we must fail because we cannot trust this stack).
\item To ensure that the check is made, we require it in the semantic condition (at this moment of time not defined, so we are yet to see what it looks like).
\end{itemize}

>>>>

\dominique{27-09-2017: I think we need to do something about tail calls in the source language?}
\lau{27-09-2017: In what sense? Do you want to allow them (I do not think they are allowed at the moment?)}

@@

The $\decInstr{}$ function must be surjective and injective for all non-$\tfail$ instructions. The $\encInstr{}$ function must be injective. Further $\decInstr{}$ must be the left inverse of $\encInstr{}$ that is for all $i \in \Instr$
\[
  \decInstr{}(\encInstr{i}) = i
\]

@@

\paragraph{Call implementation}
\label{par:call-impl}
\lau{16-10-2017: Fill this paragraph with $\scall{}{}$ implementation, i.e., the instructions $\scall[1]{r_1}{r_2} \dots \scall[\calllen]{r_1}{r_2}$ is equal to. Add convenience notation like $\update{\mem.a}{\scall{r_1}{r_2}} = \dots$}
\lau{16-10-2017: Implementation of $\scall{}{}$ may need registers for temporary registers - we should update the semantics accordingly.}
\[
  \begin{array}{l}
    \tgete{r_{1t}}{\rstk}\\
    \tstore{\rstk}{r_{t1}}\\
    \tcca{\rstk}{-1}\\
    \tgeta{r_{t1}}{\rstk}\\
    \tsplit{\rstk}{\rretd}{\rstk}{r_{t1}}\\
    \tmove{r_{t1}}{\pcreg}\\
    \tsetatob{r_{t1}}{r_{t1}}\\
    \tload{r_{t1}}{r_{t1}}\\
    \tcca{r_{t1}}{{\color{red} \dots}}\\
    \tcseal{\rretd}{\rretd}{r_{t1}}\\
    \tmove{\rretc}{\pcreg}\\
    \tcca{\rretc}{5} \text{ //magic number is offset to return code}\\
    \tcseal{\rretc}{\rretc}{r_{t1}}\\
    \tmove{r_{t1}}{0}\\
    \txjmp{r_1}{r_2}\\
    \text{// the following is the return code}\\
    \tgetb{r_{t1}}{\rstk}\\
    \tminus{r_{t1}}{t_{t1}}{\stkb} \text{ //$\stkb$ is the stack base constant} \\
    \tmove{r_{t2}}{\pcreg}\\
    \tcca{r_{t2}}{5} \text{ //magic number is the offset to fail} \\
    \tjnz{r_{t2}}{r_{t1}} \\
    \tcca{r_{t2}}{1} \text{ //magic number is the offset to after fail} \\
    \tjmp{r_{t2}} \\
    \tfail \\
    \tsplice{\rstk}{\rstk}{\rdata} \\
    \tcca{\rstk}{1}\\
    \tmove{r_{t1}}{0}\\
    \tmove{r_{t2}}{0}
  \end{array}
\]
{\color{red} \dots} needs to be the offset to the seal we want to use.
The call code does the following:
\begin{itemize}
\item Store the end address of the stack to the stack (this ensures the stack is non-empty), and decrement the pointer according to convention.
\item Get the current address of the stack pointer and split according to it.
\item Retrieve the seal of the program.
\item cca the seal, so the seal to be used is active 
\item seal our private part of the stack capability.
\item Move the pc out of the pc register, adjust it to point to the first address of the return code, and seal it.
\item clear the temporary register.
\item cross jump to the two specified registers.
\item Upon return:
  \begin{itemize}
  \item get the base of the stack and check that it matches up with the global base of the stack.If not, fail.
  \item splice the returned stack pointer and the stack pointer for our private stack.
  \item adjust stack pointer to first empty address (the address with the end address is considered free)  
  \item clear the temporary registers.
  \end{itemize}
\end{itemize}

@@

\begin{definition}
  \label{def:wxorx}
  ~\\
  A word $w$ is said to respect write-xor-execute iff
  \begin{itemize}
  \item $w \in \ints$
  \item $w = \seal{\_,\_,\_,\_}$
  \item $w = ((\perm,\_),\_,\_,\_) \tand \perm \neq \rwx$
  \item $\seal{\_,\vsc}$ and $\vsc$ respects write-xor-execute
  \item $\src{\stkptr{\perm} \tand \perm \neq \rwx}$
  \item $\src{w = \retptrc(\_,\_)}$
  \item $\src{w = \retptrd(\_,\_)}$
  \end{itemize}
~\\
  A register file $\reg$ is said to respect write-xor-execute with slack $[\bar{r}]$ iff \\
  For all $r \in \RegName$ $\reg(r) \setminus [\bar{r}]$ respects write-xor-execute.
~\\
~\\
  A memory segment $\ms$ is said to respect write-xor-execute iff\\
  For all $\aaddr \in \dom(\ms)$ $\ms(\aaddr)$ respects write-xor-execute
\end{definition}

@@

\section{Logical Relation}
\subsection{Worlds}
\lau{30-10-2017: TODO: solve recursiveness in domain equation in the following. }
\lau{02-11-2017: We may need revoked regions again to handle the stack.}
\lau{06-11-2017: If we need to introduce revoked regions like we had for the local capabilities, we may consider the following: For the future world relation require the existence of a bijection from region names to region names. For all regions in the present world the future region is given by the bijection. This would mean that reinstantiation would happen simply by giving a bijection that maps to the new region. In other words, the weird shadowing would not be necessary.
  \begin{mathpar}
    \inferrule{ \dom(W') \supseteq \dom(W) \\ \exists m : \RegionName \fun \RegionName, \text{ $m$ injective} \ldotp \forall r \in \dom(W) \ldotp W'(m(r)) \pubft W(r) }
    { W' \pubft W }
  \end{mathpar}
}
\[
  \World = \RegionName \parfun \Region
\]
where $\RegionName = \nats$.
\[
  \Region = \left\{
  \begin{array}{l}
    \{\spatial \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})\\
    \{\spatialo \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})\\
    \{\pure \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})
  \end{array} \right.
\]
where $\spatial$ and $\spatialo$ are regions governing segments of memory addressed by linear capabilities. $\spatialo$ signifies that this region is addressable. $\spatial$ signifies that the region is not owned and can thus not be addressed. Finally, $\pure$ signifies that the region is only addressed by non-linear capabilities. Notice that no region allows for both linear and non-linear capabilities to address it.
\lau{26-20-2017: We may also want to build the infrastructure for the stack more explicitly into the worlds (but maybe we can handle it in the memory satisfaction).}
\lau{26-20-2017: It is not clear to me what the monotonicity requirements on the memory interpretation functions should be.}
  \[
   \Rels = \{\phi_\pub \times \phi \in \powerset{\State^2}\times \powerset{\State^2} \mid \phi_\pub, \phi \text{ is reflexive and transitive and } \phi_\pub \subseteq \phi \}
  \]

Disjoint union of worlds. Joins together two alike worlds with strictly disjoint ownership over $\spatialo$-regions.
\[
  W_1 \oplus W_2 = W
  \begin{array}[t]{l}
    \text{if $\dom(W) = \dom(W_1) = \dom(W_2) \tand$} \\
    \forall r \in \dom(W) \ldotp W(r) = W_1(r) \oplus W_2(r)
  \end{array}
\]
$\oplus$ is defined as follows
\begin{align*}
  (\pure,\var{sm}) \oplus (\pure,\var{sm}) =  & \; (\pure,\var{sm}) \\
  (\spatialo,\var{sm}) \oplus (\spatial,\var{sm}) = & \; (\spatial,\var{sm}) \oplus (\spatialo,\var{sm})\\
                                           =  & \; (\spatialo,\var{sm}) \\
  (\spatial,\var{sm}) \oplus (\spatial,\var{sm}) =  & \; (\spatial,\var{sm}) \\
\end{align*}
and for all other cases $\oplus$ is undefined. Specifically, $\oplus$ is not defined when both sides are a $\spatialo$-region. It is further not defined if the two sides do not agree on region type and state machine. Here $\var{sm}$ is the state machine of the region.

@@

\subsection{Memory satisfaction}
\[
  \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W} \text{ iff } 
  \left\{
    \begin{array}{l}
      \exists R_\ms : \dom(W) \fun \MemSeg^2 \ldotp \\
      \quad \ms_T = \biguplus_{r \in \dom(W)} \pi_2(R_\ms(r)) \wedge \\
      \quad \stk = (\_,\ms_{\stk,0}) :: \dots :: (\_,\ms_{\stk,m}) \wedge \\
      \quad \exists r_0 , \dots, r_m, r_{m+1} \ldotp \pi_1(R_\ms(r_0,\dots,r_m)) = \ms_{\stk,0},\dots , \ms_{\stk,m} \wedge \\
      \qquad \pi_1(R_\ms(r_{m+1})) = \ms_\stk \wedge \\
      \qquad \ms_S \uplus \ms_\stk \uplus \ms_{\stk,0} \uplus \dots \ms_{\stk,m} = \biguplus_{r \in \dom(W)} \pi_1(R_\ms(r)) \wedge \\
      \qquad \exists R_w : \dom(W) \fun \World \ldotp\\
      \quad \qquad W = \bigoplus_{r \in \dom(W)} R_w(r) \wedge \\
      \quad \qquad \forall r \in \dom(W) , n' < n \ldotp \\
      \qquad \qquad \npair[n']{R_\ms(r)} \in W(r).H \; (W(r).s) \; (\xi^{-1} (R_w (r))) 
    \end{array}
  \right.
\]
In the definition of memory satisfaction, $R_\ms$ acts as a partitioning in the two related memories. As the source language is a special interpretation of the target language, part of the target memory needs to relate to the stack in the source language. We also need to split up the world which is done by $R_W$.We split up the world in order to make sure that the linearity constraints imposed by the world are satisfied.
\lau{30-10-2017: It is not clear to me whether this definition should also talk about $\opc$.}

\subsection{Relation}
\lau{19-10-2017: Rest of source configuration needs to come from somewhere in $\lre$.}
\lau{19-10-2017: Maybe linearity should be taken care of in just one place (instead of in $\lrr$ relation and mem satisfaction). That is, add a predicate that checks whether the configuration satisfies linearity (such a condition may be to syntactic to work?).}
In the following definitions, \src{blue} is used to indicate values related to the source machine. This is unlike previous definitions, where \src{blue} is used  indicate source language specific parts of definitions.
\[
  \lre(W) = \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
    \begin{array}{l}
      \forall n' \leq n, \src{k_{d,S}}, \src{k_{c,S}}, k_{d,T}, k_{c,T}, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
      \quad \forall W_\lrk, W_\lrr, W_\lrm \ldotp \\
      \qquad\npair[n']{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \in \lrp(W) \wedge \\
      \qquad\npair[n']{\stpair[.]{k_{d,S},k_{c,S}}{k_{d,T},k_{c,T}}} \in \lrk(W_\lrk) \wedge \\
      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
      \qquad\Rightarrow\\
      \qquad\npair[n']{\left(\array{l}\src{(\ms_S,\reg_S\updReg{\pcreg,\rdata,\rretc,\rretd}{v_{c,S},v_{d,S},k_{c,S},k_{d,S}},\stk, \ms_\stk)},\\
                                          (\ms_T,\reg_T\updReg{\pcreg,\rdata,\rretc,\rretd}{v_{c,T},v_{d
,T},k_{c,T},k_{d,T}})\endarray\right)}
      \in \lro(W\oplus W_\lrk \oplus W_\lrr \oplus W_\lrm)
    \end{array}
    \right\}
\]
\lau{24-10-2017: Dominique, I do not see how the splitting of worlds is going to work out. I can see how it will enforce linearity, but is you give a "spatial" region (the kind of region that is only allowed to go to one part of a split) to the $\lrr$-relation, then the memory satisfaction does not have it, and it cannot find a memory segment for this part. In other words, the memory satisfaction will not be able to find a concrete memory segment for segments governed by a linear capability in the register file. Maybe the memory satisfaction should have all of the parts, so it can represent them (but also limit what linear capabilities can be used).}

@@

\lau{03-11-2017: TODO take a look at linearity of this definition - in particular the recovered pc and stack pointer.}
\[
  \lrk(W) = \left\{ \npair{\left(\array{l}\src{\retptrd(\sigma,(\aaddr_d,\eaddr_d))},\\\src{\retptrc(\sigma,(\perm,\lin,\baddr_c,\eaddr_c,\aaddr_c))},\\k_{d,T},k_{c,T}\endarray\right)} \middle|
    \begin{array}{l}
      k_{d,T} = \sealed{\sigma,c_d} \wedge\\
      k_{c,T} = \sealed{\sigma,c_c} \wedge\\
      c_c = ((\perm,\lin),\baddr_c,\eaddr_c,\aaddr_c) \wedge\\
      c_d = ((\rw,\linear),\aaddr_d,\eaddr_d,\aaddr_d-1) \wedge\\
      \forall n' \leq n, W' \pubft W \ldotp \\
      \quad \forall, W_\lrr, W_\lrm \ldotp  \\
      \qquad\opc = c_c \wedge\\
      \qquad\dom(\ms_{\stk,f}) = [\aaddr_d,\eaddr_d] \wedge\\
      \qquad\npair[n']{\src{\retptrd(\sigma,\_)},\src{\retptrc(\sigma,\_)},k_{d,T},k_{c,T}} \in \lrp(W') \wedge \\
      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk::(\opc,\ms_{\stk,f}),\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
      \qquad\src{reg_S}(\rstk) = \src{\stkptr{\rw,\baddr_\stk,\eaddr_\stk,\_}} \wedge \\
      \qquad\eaddr_\stk = \aaddr_d-1\\
      \qquad\Rightarrow\\
      \qquad\npair[n']{\left(\array{l}\src{\left(\arraycolsep=0pt\array{l}\ms_S,\reg_S\updReg{\pcreg,\rstk}{\opc,\stkptr{\rw,\baddr_\stk,\eaddr_d,\aaddr_d}},\\\stk, \ms_\stk \uplus \ms_{\stk,f}\endarray\right)},\\
                                          (\ms_T,\reg_T\updReg{\pcreg,\rstk}{\opc,((\rw,\lin),\baddr_\stk,\eaddr_d,\aaddr_d)})\endarray\right)}
      \in \lro(W' \oplus W_\lrr \oplus W_\lrm)
    \end{array}
  \right\}
\]

@@

\[
  \lrr(W) = \left\{ \npair{\stpair{\reg}{\reg}} \middle|
    \begin{array}{l}
      \exists S : \RegName \fun \World \ldotp \\
      \quad S \text{ is a separation of $W$} \tand\\
      \quad \forall r \in \RegName \setminus \{\pcreg, \rretd,\rretc,\rdata \}\ldotp\\
      \qquad\npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrv(S(r))
    \end{array}
            \right\}
\]
Where $S\text{ is a separation of $W$}$ means that no two partition share the same ``spatial'' region (whatever that is).

@@

One way to solve this problem is to make the "top address" of the stack known. There are many ways to do this, but we have chosen to do the following:
\begin{itemize}
\item The stack grows downwards, so the ``last address'' of the stack is the base address of the initial stack capability.
\item The base address of the stack capability is a fixed address, so in the semantics, it will be expressed as a constant that is publicly known.
\item At some point before a call, it must be checked whether the stack we are using actually has the globally known base address (if not we must fail because we cannot trust this stack).
\item To ensure that the check is made, we require it in the semantic condition (at this moment of time not defined, so we are yet to see what it looks like).
\end{itemize}
